Ages of Planets and Stars Observed by Kepler

### 2024-04-17 paper/ms.tex
update ms

% <250 words
\begin{abstract}
  Recent studies of stars in open clusters have yielded a detailed picture
  of how the rotation rates and lithium content of FGK stars evolve
  during the main sequence.
  %
  Here we apply this knowledge to planets and stars observed by the
  primary Kepler mission (2009-2013), and report the resulting rotation and
  lithium-based ages.
  %
  Rotation periods are collected from previous analyses of
  Kepler data; lithium is measured from new and archival HIRES
  spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is likely valid. %, and for many more for which the underlying assumptions of
%  the technique may be compromised.
  %
  The relative statistical precisions from rotation, $\sigma_{t_{\rm
  gyro}} / t_{\rm gyro}$, are XX +YY -ZZ\%, which is X.X +Y.Y -Z.Z more
  statistically precise than isochronal ages for the same stars, and
  $\approx$X.X times less precise than cluster-based ages.
  %
  The results include \nplyounggyrotwosigma\
  planets younger than 1\,Gyr at 2$\sigma$ (\nplyounggyrothreesigma\
  planets at 3$\sigma$).
  %
  While lithium provides only marginal added information for most of
  this sample,
  we report robust two-sided age measurements from both rotation and
  lithium for 14 confirmed $<$500\,Myr planets, including five
  mini-Neptunes, five super-Earths, three Earths and sub-Earths, and a
  Neptune-sized planet.
  %
  In the broader Kepler target star sample,
  we find
  that the gyrochrone age distribution rises linearly from 0 to
  3\,Gyr, with stars 2\,Gyr old being twice as common in the Kepler
  field as stars 1\,Gyr old.
  %
  This approach offers greater sensitivity to main sequence ages than
  previous analyses of the Milky Way's star formation history, and the
  resulting age distribution bounds the relative balance of ``young''
  and ``old'' planets detectable not only by Kepler, but also by
  future wide-field transit surveys.
  %
%  The sample of detected Kepler planets has a similar age
%  distribution, though it is slightly deficient at $\lesssim$0.2\,Gyr,
%  likely due to decreased detection sensitivity.
  %
  %Combining our stellar age catalog with Kepler's completeness products, we perform
  %a hierarchical bayesian analysis to infer the average number of planets per star
  %as a function of both stellar and planetary parameters.
  %
  %While we recover many previously known correlations with stellar mass and metallicity,
  %our analysis also highlights the importance of age in shaping the ``sub-Saturn desert'',
  %and in influencing the evolution of the mini-Neptune and super-Earth populations.
  %TODO will this be the case?
%  We also comment on a few previously reported evolutionary trends in
%  the size evolution of close-in mini-Neptunes (2-4\,\rearth) and
%  super-Earths (1-2\,\rearth), a few of which we recover at comparable
%  ($\sim$3$\sigma$) significance to previous work, and one of which we
%  note for the first time.
\end{abstract}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_selsnrupper3_strictyoung.pdf}
		}

		\subfloat{
			\includegraphics[width=0.8\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung.pdf}
		}
	\end{center}
	\vspace{-0.5cm}

@@

\startlongtable
\begin{deluxetable*}{lllllllrrrrr}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
	\tablecaption{Young planets and planet candidates.
		%The non-truncated machine readable versions are accessible both
		%through the online journal, and at
		%\url{https://zenodo.org/record/8327508}.
		\label{tab:youngplanets}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ & $t_{\rm gyro}$ & $t_{\rm Li} $ & $R_{\rm p}$ & $P_{\rm orb}$ & $f_{\rm RUWE}$ & $f_{\rm grazing}$ & Spec? \\
  -- &   -- & K & days & m\AA & Myr & Myr &    Earths &    days &      bool &   bool  & bool \\
	\hline
	\input{table_subgyr_confirmed} %\\
	%\hline
	\input{table_subgyr_candidate}
	\enddata
	\tablecomments{This table includes X, Y, Z... The machine-readable version,
		available online, includes additional columns for .... 
	}
\end{deluxetable*}
%\endlongtable

>>>>

% <250 words
\begin{abstract}
  Recent studies of stars in open clusters have yielded a detailed picture
  of how the rotation rates and lithium content of FGK stars evolve
  during the main sequence.
  %
  Here we apply this knowledge to planets and stars observed by the
  primary Kepler mission, and report the resulting rotation and
  lithium-based ages.
  %
  Rotation periods are collected from previous analyses of
  Kepler data; lithium is measured from new and archival HIRES
  spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is likely valid. %, and for many more for which the underlying assumptions of
%  the technique may be compromised.
  %
  The relative statistical precisions from rotation, $\sigma_{t_{\rm
  gyro}} / t_{\rm gyro}$, are XX +YY -ZZ\%, which is X.X +Y.Y -Z.Z more
  statistically precise than isochronal ages for the same stars, and
  $\approx$X.X times less precise than cluster-based ages.
  %
  The results include \nplyounggyrotwosigma\
  planets younger than 1\,Gyr at 2$\sigma$ (\nplyounggyrothreesigma\
  planets at 3$\sigma$).
  %
  While lithium provides only marginal added information for most of
  this sample,
  we report robust two-sided age measurements from both rotation and
  lithium for 14 confirmed $<$500\,Myr planets, including five
  mini-Neptunes, five super-Earths, three Earths and sub-Earths, and a
  Neptune-sized planet.
  %
  In the broader Kepler target star sample,
  we find
  that the gyrochrone age distribution rises linearly from 0 to
  3\,Gyr, with stars 2\,Gyr old being twice as common in the Kepler
  field as stars 1\,Gyr old.
  %
  This approach offers greater sensitivity to main sequence ages than
  previous analyses of the Milky Way's star formation history, and the
  resulting age distribution bounds the relative balance of ``young''
  and ``old'' planets detectable not only by Kepler, but also by
  future wide-field transit surveys.
  %
%  The sample of detected Kepler planets has a similar age
%  distribution, though it is slightly deficient at $\lesssim$0.2\,Gyr,
%  likely due to decreased detection sensitivity.
  %
  %Combining our stellar age catalog with Kepler's completeness products, we perform
  %a hierarchical bayesian analysis to infer the average number of planets per star
  %as a function of both stellar and planetary parameters.
  %
  %While we recover many previously known correlations with stellar mass and metallicity,
  %our analysis also highlights the importance of age in shaping the ``sub-Saturn desert'',
  %and in influencing the evolution of the mini-Neptune and super-Earth populations.
  %TODO will this be the case?
%  We also comment on a few previously reported evolutionary trends in
%  the size evolution of close-in mini-Neptunes (2-4\,\rearth) and
%  super-Earths (1-2\,\rearth), a few of which we recover at comparable
%  ($\sim$3$\sigma$) significance to previous work, and one of which we
%  note for the first time.
\end{abstract}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.78\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_selsnrupper3_strictyoung.pdf}
		}

		\subfloat{
			\includegraphics[width=0.78\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung.pdf}
		}
	\end{center}
	\vspace{-0.5cm}

@@

\startlongtable
\begin{deluxetable*}{lllllllrrrrrr}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
	\tablecaption{Young planets and planet candidates.
		%The non-truncated machine readable versions are accessible both
		%through the online journal, and at
		%\url{https://zenodo.org/record/8327508}.
		\label{tab:youngplanets}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
  KOI & Kepler &  $T_{\rm eff}$ & $P_{\rm rot}$ & EW$_{\rm Li}$ & $t_{\rm gyro}$ & $t_{\rm Li} $ & $R_{\rm p}$ & $P_{\rm orb}$ & $f_{\rm RUWE}$ & $f_{\rm grazing}$ & $Q_{\rm gyro}$ &  Spec? \\
  -- &   -- & K & days & m\AA & Myr & Myr &    Earths &    days &      bool &   bool  & int & bool \\
	\hline
	\input{table_subgyr_confirmed} %\\
	%\hline
	\input{table_subgyr_candidate}
	\enddata
	\tablecomments{This table includes X, Y, Z... The machine-readable version,
		available online, includes additional columns for .... 
	}
\end{deluxetable*}
%\endlongtable
